---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  eval = FALSE
)
```

# epishiny

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

Interactive web-based data visualisations and dashboards are an increasingly common method 
of monitoring infectious disease outbreaks. Whilst R is now a commonly used tool for analysis 
and data visualisation in epidemiology, epidemiologists will often lack the knowledge required 
to be able to produce interactive dashboards directly from within their R workflows, 
resulting in external propriatory software being used instead.

`epishiny` aims to bridge this gap by providing simple functions that produce engaging, 
feature-rich interactive visualisations and dashboards from epidemiological linelist data 
using R's [`shiny`](https://shiny.posit.co/) web-framework.

## Installation

You can install the development version of epishiny from GitHub with either the [remotes](https://remotes.r-lib.org/) or [pak](https://pak.r-lib.org/) package:

``` r
# install.packages("remotes")
remotes::install_github("epicentre-msf/epishiny")

# install.packages("pak")
pak::pkg_install("epicentre-msf/epishiny")
```

# Example

```{r eval=TRUE, file='inst/examples/doc-examples/setup.R'}
```
