---
title: "Contributing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Contributing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`epishiny` welcomes new module contributions to the package as well as feature improvements to existing modules.
Contributors will require a solid understanding of how the R 
[`shiny`](https://shiny.posit.co/r/getstarted/shiny-basics/lesson1/index.html) web-framework works and the ability to program with 
reactive objects, as well as knowledge of how to organise your code as a composable shiny 'module'. 
See the [Mastering Shiny book](https://mastering-shiny.org/index.html) and the 
[modules chapter](https://mastering-shiny.org/index.html) to learn more. 

An understanding of how an R package is structured is also 
required and how to prepare functions for exportation with appropriate documentation using the 
[`roxygen2`](https://cran.r-project.org/web/packages/roxygen2/vignettes/roxygen2.html) package.

## Workflow 

Contributions should be made as pull-requests on github from your own fork of the source package, 
which will then be reviewed by package maintainers.

## Package Structure

Below is the current package structure at 2 levels of recursion:

```{r echo=FALSE}
fs::dir_tree(here::here(), recurse = 2)
```

The important directory is the `R` directory. This is the source code of the package and where any code contributions must live. 
To contribute a new module, create a new R script inside this directory with a number and concise, informative name. Add your module 
UI and server functions to this script as well as any helper functions specific to this module. More general utility functions can 
be added to the `R/utils.R` script.

## Standards 

New modules will need to meet certain criteria to be able to work in harmony with the other modules in the package. 

### User Interface

The primary requirement is to use UI components from the [`bslib` package](https://rstudio.github.io/bslib/index.html). `bslib` 
provides a modern UI toolkit for Shiny and R Markdown based on Bootstrap. Default shiny UI uses an old version of Bootstrap (v3) but 
`bslib` provides access to newer versions (latest currently v5). The package is maintained by Posit, the authors of shiny, and is now their
recommended shiny UI framework. See the [package website](https://rstudio.github.io/bslib/index.html) for documentation on all of the available components.

If your module will produce a single graphic or table output, it is recommended to use a `bslib::card()` as the module UI wrapper. If you need 
multiple tabs within the module use `bslib::navset_card_tab()`. See the package's [card article](https://rstudio.github.io/bslib/articles/cards/index.html#multiple-tabs) 
for more details and features.

**Don't forget to wrap any input IDs in your module UI in the `ns()` namespace function. See an existing module UI function for examples.**

### Server

There are no specific requirements for the module server code to meet other than to try to be as efficienct and fast as possible.

### Outputs

The package currently uses the [highcharter](https://jkunst.com/highcharter/index.html) package for interactive graphics, 
[leaflet](https://rstudio.github.io/leaflet/) for interactive maps and [reactable](https://glin.github.io/reactable/index.html) and 
[gtsummary](https://www.danieldsjoberg.com/gtsummary/) packages for tables. It is recommended to use one of these packages for your outputs if possible to maintain style and functionality
across modules and reduce the number of dependencies (more below).

### Code style

Please follow the [Tidyverse style guide](https://style.tidyverse.org/index.html) as much as possible. Use the [style](https://styler.r-lib.org/) 
package to restyle code before submitting a pull-request.

### Dependencies

Due to the large scope of this package the number of dependencies is already large with multiple htmlwidget libraries being used.
Efforts should be made to keep only include essential dependencies in your contribution. For example, if you import a library to use a 
single simple function, it is recommended to code this as your own utility function within the package and avoid importing the external package.