---
title: "Using epishiny with aggregated data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Epidemiological surveillance data is often reported in an aggregated form by ministries of health.
A typical aggregation would be by a geographical and time unit, cases and deaths by health areaa and week for example.

You can use `epishiny` to visualise data in this form by declaring one or more `count_vars` in the data 
(numeric columns containing the aggregation count totals).

Let's run through an example using [WHO's COVID-19 daily cases and deaths dataset](https://covid19.who.int/data).
The data contains daily case and death totals per country, so we can visualise both the time and place component 
using `epishiny`.

## Load libraries

```{r setup}
suppressPackageStartupMessages(library(readr))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(rnaturalearth))
suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(epishiny))
```

## Import aggregated COVID-19 data from WHO

```{r covid-data}
df_who_covid <- read_csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")

glimpse(df_who_covid)
```

## Import country boundaries with rnaturalearth

```{r geo-data}
world_map <- ne_countries(scale = "small", type = "countries", returnclass = "sf") %>%
  st_transform(crs = 4326) %>%
  select(iso_a2, name, pop_est)

# setup the geo layer for epishiny
geo_data <- geo_layer(
  layer_name = "Country",
  sf = world_map,
  name_var = "name",
  pop_var = "pop_est",
  join_by = c("iso_a2" = "Country_code")
)

geo_data
```

## Define count variables in the data 

We are only insterested in the new case and death variables, since the time module will handle calculating
cumulative numbers for us. Here we supply a named vector to show different variable labels in the module's
indicator select input.

```{r count-vars}
count_vars <- c("Cases" = "New_cases", "Deaths" = "New_deaths")
```

## Launch time module

```{r time-module, eval=FALSE}
launch_module(
  module = "time",
  df = df_who_covid,
  date_vars = "Date_reported",
  group_vars = "WHO_region",
  count_vars = count_vars,
  show_ratio = TRUE,
  ratio_lab = "CFR",
  ratio_numer = "New_deaths",
  ratio_denom = "New_cases"
)
```

## Launch place module

```{r place-module, eval=FALSE}
# filter to data in ongoing here for more relevant attack rate estimates
map_data_filter <- df_who_covid %>% 
  filter(between(Date_reported, as.Date("2023-01-01"), as.Date("2023-12-31")))

launch_module(
  module = "place",
  df = map_data_filter,
  geo_data = geo_data,
  count_vars = count_vars
)
```

## Launch person module

Now the COVID-19 data has no age or sex variables so we can't use the person module, but for demonstation
purposes we will show this can also be used with an aggregated data set below.

```{r}
# create a data set with case and death counts aggregated by age group and sex
age_levels <- c("<5", "5-17", "18-24", "25-34", "35-49", "50+")
sex_levels <- c("Male", "Female")

df_as <- tibble(
  sex = factor(c(rep(sex_levels[1], 6), rep(sex_levels[2], 6))),
  age_group = factor(rep(age_levels, 2), levels = age_levels),
  cases = round(runif(12, 20, 100)),
  deaths = round(runif(12, 0, 20)),
)

df_as
```

```{r eval=FALSE}
# launch the module passing age, sex and count_var info
launch_module(
  module = "person",
  df = df_as,
  age_group_var = "age_group",
  sex_var = "sex",
  male_level = "Male",
  female_level = "Female",
  count_vars = c("cases", "deaths")
)
```
