---
title: "Using epishiny with linelist data"
author: "Sebastian Funk"
bibliography: references.bib
nocite: '@*'
link-citations: true
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using epishiny with linelist data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library("knitr")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The _epishiny_ package can work with pre-aggregated data or with line lists, i.e. data sets in the form of 
tables that contain cases as one line per individual. 

The package provides a range of visualisations of different aggregations of the data that can either 
be launched as individual modules or as part of a shiny dashboard that can be run locally or deployed on a server.

In this demonstation we will walk throught the steps of preparing external data then visualise it by 
launching individual modules from within an R script. 

# Load in data

The package comes with a built in example line list `df_ll`, but a user can also bring their own data.
Here we will use a line list of Ebola in Sierra Leone published in @Fang2016.

```{r load_data}
suppressPackageStartupMessages(library("readr"))
suppressPackageStartupMessages(library("dplyr"))
suppressPackageStartupMessages(library("purrr"))
suppressPackageStartupMessages(library("sf"))
suppressPackageStartupMessages(library("epishiny"))

url <- paste(
  "https://raw.githubusercontent.com/parksw3/epidist-paper/main/data-raw/",
  "pnas.1518587113.sd02.csv",
  sep = "/"
)
df <- read_csv(url)
glimpse(df)
```

```{r load_geo_data_hdx, eval=FALSE, include=FALSE}
# this link isn't working anymore so switched to geoBoundaries below
## common element of both shapefile URLs
hdx_dir <-
  "https://data.humdata.org/dataset/a4816317-a913-4619-b1e9-d89e21c056b4"
## shapefile names and resource ID (from URL)
shapefiles <- list(
  adm2 = list(
    filename = "sle_admbnda_adm2_1m_gov_ocha.zip",
    resource = "b3963917-8550-478d-9363-736492bf209a"
  ),
  adm3 = list(
    filename = "sle_admbnda_adm3_1m_gov_ocha_20161017.zip",
    resource = "e2aa661d-af2f-42d8-bdea-c7e16a00bdb2"
  )
)
## create temporary dir for downloading
tmpdir <- tempdir()

## download and load shapefiles
shapes <- map(shapefiles, \(x) {
  ## construct URL
  url <- paste(
    hdx_dir, "resource", x$resource, "download", x$filename, sep = "/"
  )
  ## construct file name
  ## (`read_sf` expects ending `.shp.zip` for zipped shapefiles)
  destfile <- sub("\\.zip$", ".shp.zip", file.path(tmpdir, x$filename))
  download.file(url, destfile = destfile)
  return(read_sf(destfile))
})

# little bit of cleaning on district names to match with data
shapes$adm2$shapeName <- gsub("Area ", "", shapes$adm2$shapeName)

map(shapes, head)
```

# Set up geo data

We next need geological data if we want to show maps.

[geoBoundaries](https://www.geoboundaries.org/) host sub-national administrative boundary 
shapefiles in various formats for every country in the world on github.

The line list contains Districts (admin 2 level) and Chiefdoms (admin 3 level).
We can download the corresponding data sets from geoBoundaries via the following URLs.

```{r load_geo_data_hdx}
shape_paths <- list(
  adm2 = "https://github.com/wmgeolab/geoBoundaries/raw/905b0ba/releaseData/gbHumanitarian/SLE/ADM2/geoBoundaries-SLE-ADM2_simplified.geojson",
  adm3 = "https://github.com/wmgeolab/geoBoundaries/raw/905b0ba/releaseData/gbHumanitarian/SLE/ADM3/geoBoundaries-SLE-ADM3_simplified.geojson"
)

shapes <- map(shape_paths, st_read)

# little bit of cleaning on district names to match with data
shapes$adm2$shapeName <- gsub("Area ", "", shapes$adm2$admin2Name)

map(shapes, head)
```

# Launch Modules

Before we launch the modules we can define some grouping variables. If passed to the `time` or `place` modules,
a select input will appear in the 'options' dropdown allowing you to group the data by the variables you select.

If you want a more readable variable label to appear in the module rather than the variable name itself, pass a 
named vector where the name is the label and the value is the variable name. 

In our example we use sex and district as a variable, but since they are already formatted as labels there is 
no need to pass names:

```{r group_vars}
group_vars <- c("Sex", "District")
```

## Place module

Now that we have the shapefiles we can collate the information contained in the format that _epishiny_ expects
using the `geo_layer()` function. Since we are using more than one geo layer, we combine them in a list:

```{r geo_data, eval=FALSE}
geo_data <- list(
  geo_layer(
    layer_name = "District",
    sf = shapes$adm2,
    name_var = "shapeName",
    join_by = c("shapeName" = "District")
  ),
  geo_layer(
    layer_name = "Chiefdom",
    sf = shapes$adm3,
    name_var = "shapeName",
    join_by = c("shapeName" = "Chiefdom")
  )
)
```

We use this to launch the place module:

```{r place_module, eval = FALSE}
launch_module(
  module = "place",
  df = df,
  geo_data = geo_data,
  group_vars = group_vars[1] # only pass sex variable since district is already visualised on map
)
```

## Time module

To launch the time module, we need to pass the date variable(s) in the line list we want to use for the x-axis.

Notice that the date variables are of character class in the data but they are automatically parsed to date class
in the time module via the lubridate::as_date function.

```{r time_module, eval = FALSE}
launch_module(
  module = "time",
  df = df,
  date_vars = c("Date of symptom onset", "Date of sample tested"),
  group_vars = group_vars
)
```

## Person module

Finally, we can plot an age/sex pyramid using the person module, passing age and sex variable names
and the levels in the sex variable to encode male and female:

```{r person_module, eval = FALSE}
launch_module(
  module = "person",
  df = df,
  age_var = "Age",
  sex_var = "Sex",
  male_level = "M",
  female_level = "F"
)
```


# References


